Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-03 examples.tex
+

 \subsubsection{Ensuring Presence of an Entry}
After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}. Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.

@@

@@

@@

@@

@@

Regarding the realization of 
%\begin{figure}\footnotesize
%\begin{lstlisting}[mathescape]
%;;if (!entry->present) {
%  ...
%;;}
%...
%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$ 
%mov    rax,-0x18[rbp]
%;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
%mov    -0x18[rbp],rax
%movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
%add    rcx,rax
%mov    rax,-0x20[rbp]
%\end{lstlisting}
%\vspace{-1em}
%\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
%associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
%\label{fig:p2v}
%\end{figure}

>>>>

 \subsubsection{Ensuring Presence of an Entry}
After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}. Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
\begin{figure}\footnotesize
  \begin{lstlisting}[mathescape]

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{\_} \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast\sumapaces\rtv\delta }_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = None \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$
$\sqsubseteq$
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{next} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{entry\_val}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_} \ast \sumapaces\rtv\delta }_{\rtv}$  

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\_}\ast \sumapaces\rtv\delta }_{\rtv}$  

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi\cup\{\mathsf{nextpaddr}\},m)   \ast \texttt{rbp-16} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{table\_addr}  }_{\rtv}$
$\specline{\texttt{rax} \mapsto_{\textsf{v}} \textsf{nextpaddr} \ast \ulcorner \Xi !! \mathsf{nextpaddr} = Z \urcorner \ast \textsf{table\_addr} \mapsto_{\mathsf{v}} \mathsf{\mathsf{nextpaddr}} \ast \sumapaces\rtv\delta }_{\rtv}$  

@@

\mytodo{give line numbers andrelated post here and pteinitcal }

@@

Realizing page-table directory walk amounts to calling \textsf{pte\_get\_next\_table} for each level as shown in Figure \ref{walkpgdir}. Since there is not much interesting piece additional to the specification we discuss for \textsf{pte\_get\_next\_table} in Section \ref{}, in Figure \ref{}, we present the expected specification after the calls made for each level entry. At each specification Line (X,..) in Figure \ref{}, we see the accumulation of virtual-pte pointsto resources returned from \textsf{pte\_get\_next\_tabel} for each level (L4-L1), and finally return L1 level entry that is needed by \textsf{walkpgdir} at Line X in Figure \ref{}. 
;;pte_t *walkpgdir_succ(pte_t *pml4, const void *va) {
;;set up the stack for root address and virtual address    
;;pte_t *pml4_entry = &pml4[PML4EX(va)];
mov    -0x8[rbp],rsi
mov    -0x10[rbp],rdi
shr    0x27,rdi
and    0x1ff,rdi
shl    0x3,rdi
add    rdi,rsi
mov    rsi,-0x18[rbp]
;;pte_t *pdp = pte_get_next_table_succ(pml4_entry);
mov    -0x18[rbp],rdi

$\specline{}$
callq  c0 <pte_get_next_table_succ>
;;save the physical next table address in rax
mov    rax,-0x20[rbp]
$\specline{}$
;;pte_t *pdp_entry = &pdp[PDPEX(va)];
;;pte_t *pd = pte_get_next_table_succ(pdp_entry);
$\specline{}$
...
;;pte_t *pd_entry = &pd[PDEX(va)];
;;pte_t *pt = pte_get_next_table_succ(pd_entry);
 $\specline{}$
;;return &pt[PTEX(va)];
...
;; clean up the stack 
\end{lstlisting}
\vspace{-1em}
\caption{Walking page-table directory via calls to \textsf{pte\_get\_next\_table} in Figure \ref{fig:calltopteinitialize}}
\label{walkpgdir}
\end{figure}
%\begin{figure}\footnotesize
%\begin{lstlisting}[mathescape]
%;;if (!entry->present) {
%  ...
%
%;;}
%...
%$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}_{\textsf{id}}(\theta,\Xi,m)  \ast \texttt{rbp-8} \mapsto_{\textsf{v}} \textsf{entry} \ast \texttt{r8}  \mapsto_{\textsf{r}} \textsf{\_} \ast \texttt{rdi}  \mapsto_{\textsf{r}} \textsf{\_} \ast \sumpv\paddr\qfrac\true  \sumapaces\rtv\delta  }_{\rtv}$ 
%mov    rax,-0x18[rbp]
 % 
%;;uintptr_t next_virt_addr = (uintptr_t) P2V(next_phys_addr);
%mov    -0x18[rbp],rax
%movabs 0x0,rcx            187: R_X86_64_64	KERN_BASE
%add    rcx,rax
%mov    rax,-0x20[rbp]
%\end{lstlisting}
%\vspace{-1em}
%\caption{Converting a physical address, whose exsitence is made certain, of a PTE to a virtual address (w/o instruction pointer or flag updates). $R$ abbreviates the resources
%associated with \paddr\ by $\Xi$ in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}.}
%\label{fig:p2v}
%\end{figure}
